Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-05-08 manuscript/manuscript.tex
quick run through

The Kolmogorov-Smirnov (KS) test is a widely used statistical test that
assesses the conformity of a sample to a specified distribution. Its efficacy,
however, diminishes with serially dependent data and when parameters
within the hypothesized distribution are unknown. For independent data,
parametric and nonparametric bootstrap procedure are available to adjust for
estimated parameters. For serially dependent stationary data, parametric
bootstrap has been developed with a working serial dependence structure. A
counterpart for the nonparametric bootstrap approach, which needs a bias
correction, has not been studied. Addressing this gap, our study introduces a
bias correction method employing a non-parametric block bootstrap, which
approximates the distribution of the KS statistic accounting for unspecified
serial dependence and unspecified parameters. We assess its effectiveness
through simulations, scrutinizing both its size and power. The practicality of
our method is further illustrated with an examination of stock returns from the
S\&P 100 and S\&P 500 indices, showcasing its utility in real-world
applications.

@@

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines
common misapplications of the one-sample KS test. One notable inappropriate
use is when the hypothesized distribution contains unspecified parameters.
In this case, the tests are generally constructed by substituting the unknown
parameters by their estimates, and the asymptotic null distribution of
the test statistic may depend in a complex way on the unknown parameters.
The problem of unspecified parameters can be handled by a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} address this issue through a
non-parametric bootstrap method that corrects the bias in the asymptotic null
distribution.

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
The distribution of the test statistic under
the null hypothesis depends on the unknown parameters and unknown 
serial dependence. Therefore, the null distribution under dependence cannot be
used,
and the challenge is to approximate said distribution.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for,
but this could affect results if the estimated serial
dependence is too far from the truth. For this reason, a non-parametric method
is motivated.

Developing a fully non-parametric solution for the KS 
tests in the presence of serially dependent data presents a significant 
challenge. The null distribution's characteristics are intricately linked to 
the structure of the serial dependence, which can vary widely in practical 
situations. Employing a non-parametric bootstrap would necessitate defining a 
specific model for this dependence, despite the primary focus being on 
evaluating the marginal distribution of a stationary series. To date, a bias 
correction method for block bootstrap, akin to the non-parametric bootstrap
approach of \citet{babu2004goodness} for independent data,
has not been established. This study seeks 
to bridge this gap by introducing a bias correction technique for the 
non-parametric block bootstrap, tailored for use in scenarios where the 
hypothesized distribution includes unspecified parameters and the data 
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

The remainder of this paper is structured as follows: 
Section~\ref{sec:methods} provides an overview of the block bootstrap 
procedure and introduces the bias correction methodology for the 
non-parametric block bootstrap KS test. Section~\ref{sec:simu} is divided 
into two parts; initially, we evaluate the KS test's ability to maintain 
its size, i.e., its consistency in not rejecting the null hypothesis when 
it is indeed true. Subsequently, we examine the test's power, assessing 
whether it can reject the null hypothesis when it is false. Practical 
applications of our method are presented in Section~\ref{sec:real}, where 
we apply the approach to analyze if S\&P 100 and S\&P 500 index stock return
data adhere to either the Normal or the Student's $t$ distribution. The 
paper concludes with Section~\ref{sec:conclusion}, offering final thoughts 
and remarks.

@@

\subsection{Non-Parametric Bootstrap for Independent Data}

@@

\subsection{Non-Parametric Block Bootstrap for Stationary Series}

@@

Using R packages \textsl{qqplotr} and \textsl{ggplot2} \citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, the
points will be aligned with the diagonal line. This 
indicates that under the null hypothesis, our method will only reject the $H_0$
at the rate of the significance level.
\jy{Describe the table not zoomed-in figure.}\mc{addressed}
A table of the rejection rates for each scheme
also provided/

@@

From Tables~\ref{table:rr_norm} and \ref{table:rr_gamma}, we can observe 
that the agreement between empirical size (rejection rate)
and nominal size (the
significance level) improves as sample size increases, but for $\tau = -0.75$
and $\tau = 0.75$, a sample size larger than 800 seems to be necessary for that 
agreemet to be reached. Using the criteria of approximate agreement between 
empirical size and nominal size, we can say that our approach is working 
for $-0.25 < \tau < 0.25$ for $n > 100$. However a larger sample size seems 
necessary, greater than 400 perhaps, for $\tau$ as strong as $0.5$. In other
words, for weaker dependence, our method works without need for a particular
large sample size, whereas for moderate dependence, a larger sample size may 
be required. Because $\tau$ is typically less than 0.5, it is still safe to
recommend our method with a smaller sample size.

@@

Using the \textsl{tseries} package  \citep{tseries}, 
we gathered daily closing prices for the S\&P 500 index from January 1st, 2020
to December 31st, 2023. Daily returns were obtained by taking the difference in
the logarithm of the closing prices. The Kendall's $\tau$ for the series
was about -0.0298, indicating a very slight serial dependence.
Using our method, we tested that the daily returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
We tested the same hypotheses using \citet{babu2004goodness}'s 
non-parametric basic bootstrap bias correction and 
\citet{zeimbekakis2022misuses}'s parametric bootstrap method.
We used $B = 10,000$ bootstrap 
replicates in this analysis.
The p-values for these tests are highlighted in 
Table~\ref{table:SP5004}.

@@

While differences in p-values among different methods are expected, our analysis
reveals numerous instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2022misuses}.
For example, when assessing the S\&P 500 series, our method fails to reject the
hypothesis that the series follows a Student's $t$ distribution with degrees of
freedom $v = 5$ with a p-value of 0.061. In contrast, Babu and Sen's method
yields a p-value of 0.0222, and Zeimbekakis et al.'s method yields a p-value of
0.0141. Although such differences are not guaranteed, this
instance highlights a disagreement in conclusions among methods for the S\&P 500
index.

@@

The Kolmogorov-Smirnov test stands as a crucial tool for researchers across
diverse fields, enabling comparisons of population distributions. This paper
introduces a variant of the test tailored for serially dependent data without
necessitating parameter specification.
Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. Notably, it demands a larger sample
size compared to basic bootstrap methods to avoid both type I and type II 
errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust.

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with the non-parametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric
bootstrap method developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}\mc{addressed}
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA processes. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

>>>>

The Kolmogorov-Smirnov (KS) test is a widely used statistical test that
assesses the conformity of a sample to a specified distribution. Its efficacy,
however, diminishes with serially dependent data and when parameters
within the hypothesized distribution are unknown. For independent data,
parametric and nonparametric bootstrap procedure are available to adjust for
estimated parameters. For serially dependent stationary data, parametric
bootstrap has been developed with a working serial dependence structure. A
counterpart for the nonparametric bootstrap approach, which needs a bias
correction, has not been studied. Addressing this gap, our study introduces a
bias correction method employing a nonparametric block bootstrap, which
approximates the distribution of the KS statistic accounting for unspecified
serial dependence and unspecified parameters. We assess its effectiveness
through simulations, scrutinizing both its size and power. The practicality of
our method is further illustrated with an examination of stock returns from the 
S\&P 500 index, showcasing its utility in real-world
applications.

@@

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines
common misapplications of the one-sample KS test. One notable inappropriate
use is when the hypothesized distribution contains unspecified parameters.
In this case, the tests are generally constructed by substituting the unknown
parameters by their estimates, and the asymptotic null distribution of
the test statistic may depend in a complex way on the unknown parameters.
The problem of unspecified parameters can be handled by a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} address this issue through a
nonparametric bootstrap method that corrects the bias in the asymptotic null
distribution.

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
The distribution of the test statistic under
the null hypothesis depends on the unknown parameters and unknown 
serial dependence. Therefore, the null distribution under dependence cannot be
used,
and the challenge is to approximate said distribution.
\citet{zeimbekakis2022misuses} proposed a semiparametric bootstrap
where the serial dependence structure is modeled via a working serial copula.

A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semiparametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for,
but this could affect results if the estimated serial
dependence is too far from the truth. For this reason, a nonparametric method
is motivated.

Developing a fully nonparametric solution for the KS 
tests in the presence of serially dependent data presents a significant 
challenge. The null distribution's characteristics are intricately linked to 
the structure of the serial dependence, which can vary widely in practical 
situations. Employing a nonparametric bootstrap would necessitate defining a 
specific model for this dependence, despite the primary focus being on 
evaluating the marginal distribution of a stationary series. To date, a bias 
correction method for block bootstrap, akin to the nonparametric bootstrap
approach of \citet{babu2004goodness} for independent data,
has not been established. This study seeks 
to bridge this gap by introducing a bias correction technique for the 
nonparametric block bootstrap, tailored for use in scenarios where the 
hypothesized distribution has unspecified parameters and the data 
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

The remainder of this paper is structured as follows: 
Section~\ref{sec:methods} provides an overview of the block bootstrap 
procedure and introduces the bias correction methodology for the 
nonparametric block bootstrap KS test. Section~\ref{sec:simu} is divided 
into two parts; initially, we evaluate the KS test's ability to maintain 
its size, i.e., its consistency in not rejecting the null hypothesis when 
it is indeed true. Subsequently, we examine the test's power, assessing 
whether it can reject the null hypothesis when it is false. Practical 
applications of our method are presented in Section~\ref{sec:real}, where 
we apply the approach to analyze if S\&P 500 index stock return
data adheres to either the Normal or the Student's $t$ distribution. The 
paper concludes with Section~\ref{sec:conclusion}, offering final thoughts 
and remarks.

@@

\subsection{Nonparametric Bootstrap for Independent Data}

@@

\subsection{Nonparametric Block Bootstrap for Stationary Series}

@@

Using R packages \textsl{qqplotr} and \textsl{ggplot2} \citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, the
points will be aligned with the diagonal line. This 
indicates that under the null hypothesis, our method will only reject the $H_0$
at the rate of the significance level.
\jy{Describe the table not zoomed-in figure.}\mc{addressed}
A table of the rejection rates for each scheme
also provided.

@@

From Tables~\ref{table:rr_norm} and \ref{table:rr_gamma}, we can observe 
that the agreement between empirical size (rejection rate)
and nominal size (the
significance level) improves as sample size increases, but for $\tau = -0.75$
and $\tau = 0.75$, a sample size larger than 800 seems to be necessary for that 
agreemet to be reached. Using the criteria of approximate agreement between 
empirical size and nominal size, we can say that our approach is working 
for $-0.25 < \tau < 0.25$ for $n > 100$. However a larger sample size seems 
necessary, greater than 400 perhaps, for $\tau$ as strong as $0.5$. To 
summarize, for weaker dependence, our method works without need for a 
particularly
large sample size, whereas for moderate dependence, a larger sample size may 
be required. Because $\tau$ is typically less than 0.5, it is still reasonable
to
recommend our method with a smaller sample size.

@@

Using the \textsl{tseries} package  \citep{tseries}, 
we gathered daily closing prices for the S\&P 500 index from January 1st, 2020
to December 31st, 2023. Daily returns were obtained by taking the difference in
the logarithm of the closing prices. The Kendall's $\tau$ for the series
was about -0.0298, indicating a very slight serial dependence.
Using our method, we tested that the daily returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
We tested the same hypotheses using 1) the
semiparametric bootstrap method shown by \citet{zeimbekakis2022misuses} which 
accounts for both serial dependence and unspecified parameters, 2)
nonparametric basic bootstrap bias correction developed by 
\citet{babu2004goodness}, and 3) the parametric bootstrap method shown
by \citet{zeimbekakis2022misuses}.
We used $B = 10,000$ bootstrap 
replicates in this analysis.
The p-values for these tests are highlighted in 
Table~\ref{table:SP5004}.

@@

While differences in p-values among different methods are expected, our analysis
reveals numerous instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2022misuses}.
For example, when assessing the S\&P 500 series, our method fails to reject the
hypothesis that the series follows a Student's $t$ distribution with degrees of
freedom $v = 5$ with a p-value of 0.061. In contrast, 
the semiparametric method yields a p-value of 0.0192,
the basic nonparametric method
yields a p-value of 0.0222, and the parametric method yields a p-value of
0.0141. Although such differences are not guaranteed, this
instance highlights a disagreement in conclusions among methods for the S\&P 500
index.

@@

The Kolmogorov-Smirnov test stands as a crucial tool for researchers across
diverse fields, enabling comparisons of population distributions. This paper
introduces a variant of the test tailored for serially dependent data without
necessitating parameter specification.
Using simulation, we have shown that given a large enough sample of a time 
series and if the serial dependence is not too extreme, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. Notably, it demands a larger sample
size compared to basic bootstrap methods to avoid both type I and type II 
errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust.

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with the nonparametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric and semiparametric
bootstrap methods developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}\mc{addressed}
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA processes. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.
